---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message = FALSE,
  warning = FALSE,
  dev = "svg"
)
```

# indcp

<!-- badges: start -->
<!-- badges: end -->

The `indcp` package provides a set of functions for the analysis of
the individual-level child penalties (Arkhangelsky, Yanagimoto, and Zohar, 2024)

## Installation

You can install the development version of indcp from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("kazuyanagimoto/indcp")
```

## Example

This is a basic example with the simulated `base_indcp` data set:

```{r example}
library(indcp)
library(dplyr)
library(ggplot2)

base_indcp |>
  head()
```

Individual-level child penalties are estimated by `indcp()`:

```{r mdl_indcp}
mdl_base <- base_indcp |>
  filter(cage >= 25) |>
  indcp(yname = "y",
        iname = "id",
        tname = "year",
        bname = "byear",
        kname = "rel_time")

# Estimated individual-level child penalties (y_tilde)
mdl_base$df_indcp |>
  head()
```

### Aggregation of Individual-level Child Penalties

They can be aggregated to the `full`, `cage` (age at childbirth),
and `cage_byear` (age at childbirth and birth year) levels:

```{r agg_full}
summary(mdl_base) # default agg = "full"
```

```{r agg_cage}
sum_cage <- summary(mdl_base, agg = "cage")

sum_cage |>
  filter(rel_time == 0) |>
  mutate(rel_time = -1,
         mean_y_tilde = 0) |>
  bind_rows(sum_cage) |>
  filter(between(cage, 25, 34)) |>
  mutate(lbl_facet = paste0("Age ", cage)) |>
  ggplot(aes(x = rel_time, y = mean_y_tilde)) +
  geom_point() +
  geom_line() +
  geom_vline(xintercept = -1, linetype = "dashed") +
  geom_hline(yintercept = 0) +
  facet_wrap(~lbl_facet, ncol = 5) +
  labs(x = "Time to First Childbirth", y = "Child Penalties on y") +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank())
```

### Variance of Individual-level Child Penalties

Since the individual-level child penalties are estimated with measurement errors,
the variance of the `y_tilde` is not equal to the variance of
the individual-level child penalties.

The `compute_var_me` option of the `indcp` estimates the variance of
the measurement errors and the variance of the individual-level child penalties
by subtracting the variance of the measurement errors from the variance of `y_tilde`

```{r var_cage}
mdl_base_var <- base_indcp |>
  filter(cage >= 25) |>
  indcp(yname = "y",
        iname = "id",
        tname = "year",
        bname = "byear",
        kname = "rel_time",
        compute_var_me = TRUE)

sum_cage_var <- summary(mdl_base_var, agg = "cage")

sum_cage_var |>
  filter(rel_time == 0) |>
  mutate(rel_time = -1,
         sd_y_tilde_estimated = 0) |>
  bind_rows(sum_cage_var) |>
  filter(between(cage, 25, 34)) |>
  mutate(lbl_facet = paste0("Age ", cage)) |>
  ggplot(aes(x = rel_time, y = sd_y_tilde_estimated)) +
  geom_point() +
  geom_line() +
  geom_vline(xintercept = -1, linetype = "dashed") +
  geom_hline(yintercept = 0) +
  facet_wrap(~lbl_facet, ncol = 5) +
  labs(x = "Time to First Childbirth", y = "S.D. of Child Penalties") +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank())
```

