---
title: "Individual-level Child Penalties with PSID"
author: Kazuharu Yanagimoto
date: today
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  dev = "svg",
  fig.ext = "svg",
  fig.cap = ""
)
```

```{r setup}
library(unitdid)
library(dplyr)
library(ggplot2)
library(purrr)
```

### Panel Study of Income Dynamics (PSID)

The PSID data is a longitudinal household survey of the United States and one of
the longest and most famous panel surveys in the world. The study began in 1968
with a nationally representative sample of over 18,000 individuals living in
5,000 families. The data is available [here](https://psidonline.isr.umich.edu)
for the registered users.

To avoid the convoluted cleaning tasks, I mostly use the cleaning codes provided
by the [Comparative Panel File (CPF)](https://www.cpfdata.com). Their cleaning
[codes](https://www.cpfdata.com) are developed for creating a harmonized panel
data from seven countries (Australia, Germany, the UK, South Korea, Russia,
Switzerland, and the US).

To replicate this article, you need to run

1.  Run the code of the CPF only for the PSID
    -   In step 4 (Add vars constant across all waves) of the
        `CPF-Code/03_PSID/us_01_3_GEt_vars.do`, add `ER32024` (birth year of the
        first child)
    -   `us_01.dta`-`us_03.dta` will be produced
2.  Run the following code to create a dataset for the child penalties

```{r cleaning}
#| code-fold: true
#| eval: FALSE
library(haven)
cpf_us01 <- read_dta(here::here("vignettes/us_01.dta"))
cpf_us03 <- read_dta(here::here("vignettes/us_03_CPF.dta"))

ever_birth <- cpf_us01 |>
  select(pid, cyear = ER32024) |>
  filter(cyear != 9999) |>
  distinct()

cp_psid <- cpf_us03 |>
  filter(country == 3) |>
  select(id = pid,
         year = intyear,
         byear = yborn,
         gender = female,
         earn = incjobs_yg,
         hours = whweek,
         years_edu = eduy) |>
  mutate(gender = if_else(gender == 0, "Men", "Women"),
         particip = if_else(hours > 0, 1, 0),
         wage = if_else(particip == 1, earn / hours, 0),
         years_edu = if_else(years_edu < 0, NA_integer_, years_edu)) |>
  right_join(ever_birth, by = c("id" = "pid")) |>
  mutate(rel_time = year - cyear) |>
  filter(!is.na(earn))

save(cp_psid, file = here::here("vignettes/cp_psid.rds"))
```

## Analysis of Child Penalties on Participation

```{r unitdid_cage}
load(here::here("vignettes/cp_psid.rds"))

cp_psid <- cp_psid |>
  mutate(eage = cyear - byear) |>
  filter(cyear >= 1968,
         year <= 1997, # The last year of the annual data
         between(eage, 22, 40))

mdl_particip <- cp_psid |>
  unitdid(yname = "particip",
        iname = "id",
        tname = "year",
        ename = "cyear",
        bname = "byear",
        by = c("gender"),
        compute_varcov = "var")

sum_particip <- summary(mdl_particip, agg = "event_age",
                        only_full_horizon = TRUE) |>
  filter(between(event_age, 26, 32))

df_plot <- sum_particip |>
  filter(rel_time == 0) |>
  mutate(rel_time = -1, mean = 0,
         var = 0) |>
  bind_rows(sum_particip) |>
  mutate(lbl_facet = paste0("Age ", event_age))
```

### Mean of Child Penalties on Participation

```{r mean_cage}
#| code-fold: true
df_plot |>
  ggplot(aes(x = rel_time, y = mean,
             color = gender, shape = gender)) +
  geom_point() +
  geom_line() +
  geom_vline(xintercept = -1, linetype = "dashed") +
  geom_hline(yintercept = 0) +
  facet_wrap(~lbl_facet) +
  scale_color_manual(values = c("#009F8C", "#B75C9D")) +
  labs(x = "Time to First Childbirth",
       y = NULL,
       color = NULL, shape = NULL) +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        legend.position = "bottom")
```

### Standard Deviation of Child Penalties on Participation

```{r sd_cage}
#| code-fold: true
df_plot |>
  ggplot(aes(x = rel_time, y = sqrt(var),
             color = gender, shape = gender)) +
  geom_point() +
  geom_line() +
  geom_vline(xintercept = -1, linetype = "dashed") +
  geom_hline(yintercept = 0) +
  facet_wrap(~lbl_facet) +
  scale_color_manual(values = c("#009F8C", "#B75C9D")) +
  labs(x = "Time to First Childbirth",
       y = NULL, color = NULL, shape = NULL) +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        legend.position = "bottom")
```


## Scatter Plot with Covariates

To highlight the flexibility of the individual child penalties, let's plot the
relationship between the years of education and the child penalties.

```{r}
unitdid_particip <- get_unitdid(mdl_particip)
```

```{r}
#| code-fold: true
library(patchwork)

plot_scatter <- function(gender) {
  clr <- ifelse(gender == "Men", "#009F8C", "#B75C9D")

  unitdid_particip |>
    filter(gender == {{gender}}, between(rel_time, 0, 5)) |>
    mutate(lbl_facet = paste0("k = ", rel_time)) |>
    ggplot(aes(x = years_edu, y = particip_tilde)) +
    geom_point(color = clr, alpha = 0.1) +
    geom_smooth(method = "lm", color = clr, fill = clr) +
    facet_wrap(~lbl_facet, nrow = 1) +
    labs(x = "Years of Education",
         y = "CP on Participation",
         title = gender) +
    coord_cartesian(xlim = c(6, NA), ylim = c(-1.25, 0.5)) +
    theme_minimal()
}

p1 <- plot_scatter("Men")
p2 <- plot_scatter("Women")

p1 / p2
```

### On Binscatter

Another way to visualize the relationship is to use the binscatter [@cattaneo2024].

```{r}
#| include: FALSE
library(binsreg)
# Since I cannot hide plots in binsreg, I duplicate the code with `eval = FALSE`
compute_binsreg <- function(gender) {
  est <- unitdid_particip |>
    filter(gender == {{ gender }}) |>
    as.data.frame() |>
    binsreg(x = years_edu, y = particip_tilde, data = _,
            by = rel_time)

  map(0:5, ~{est$data.plot[[paste0("Group ", .x)]]$data.dots |>
               select(x, fit) |>
               mutate(rel_time = .x)}) |>
    list_rbind() |>
    mutate(gender = gender)
}

bs <- map(c("Men", "Women"), compute_binsreg) |> list_rbind()
```

```{r}
#| eval: FALSE
library(binsreg)

compute_binsreg <- function(gender) {
  est <- unitdid_particip |>
    filter(gender == {{ gender }}) |>
    as.data.frame() |>
    binsreg(x = years_edu, y = particip_tilde,
            data = _, by = rel_time)

  map(0:5, ~{est$data.plot[[paste0("Group ", .x)]]$data.dots |>
               select(x, fit) |>
               mutate(rel_time = .x)}) |>
    list_rbind() |>
    mutate(gender = gender)
}

bs <- map(c("Men", "Women"), compute_binsreg) |> list_rbind()
```

```{r}
#| code-fold: true
plot_binsreg <- function(gender) {
  clr <- ifelse(gender == "Men", "#009F8C", "#B75C9D")

  bs |>
    filter(gender == {{ gender }}) |>
    mutate(lbl_facet = paste0("k = ", rel_time)) |>
    ggplot(aes(x = x, y = fit)) +
    geom_point(color = clr) +
    facet_wrap(~lbl_facet, nrow = 1) +
    labs(x = "Years of Education",
         y = "CP on Participation",
         title = gender) +
    theme_minimal()
}

p1 <- plot_binsreg("Men")
p2 <- plot_binsreg("Women")

p1 / p2
```

## References

::: {#refs}
:::