---
title: "Data Generation Process of base_indcp"
author: Kazuharu Yanagimoto
date: '`r format(Sys.time(), "%B %d, %Y")`'
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data Generation Process of base_indcp}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  dev = "svg",
  fig.ext = "svg"
)
```

```{r setup}
library(indcp)
library(dplyr)
library(ggplot2)
library(purrr)
```

## Data Generation Process

The sample data set `base_indcp` is generated by `indcp::gen_data()` with seed `1234`.

```r
set.seed(1234)
base_indcp <- gen_data()
```

Inside of the function `gen_data()`, the following steps are performed.

### 1. Generate $i = \{1, \dots, 1000\}$ individuals

Generate $i = \{1, \dots, 1000\}$ individuals with age at child birth $a_i$
from a multinomial distribution of $f(a)$.


$$
f(a) = \begin{cases}
\frac{a - 14}{15 \cdot 16} & \text{if } a \in \{15, \dots, 29\} \\
\frac{45 - a}{15 \cdot 16} & \text{if } a \in \{30, \dots, 44\} \\
\end{cases}
$$
```{r, fig.align='center', fig.width=6}
base_indcp |>
  distinct(id, .keep_all = TRUE) |>
  mutate(event_age = cyear - byear) |>
  summarize(n = n(), .by = event_age) |>
  mutate(density = n / sum(n)) |>
  ggplot(aes(x = event_age, y = density)) +
  geom_col() +
  labs(x = "Age at First Childbirth", y = NULL) +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank())
```

### 2. Generate Individual FE $\alpha_i$ and Time FE $\lambda_t$

Generate individual fixed effects $\alpha_i$ and time fixed effects $\lambda_t$ from normal distributions.

$$
\alpha_{i} \sim \mathcal{N}(0.15, 0.1),\, \lambda_{t} \sim \mathcal{N}(0, 0.1)
$$

### 3. Generate Heterogeneous Treatment Effect $\tau_{i, k}$

For individual $i$ with age at first childbirth $a_i$, draw $\tau_{i, k} \sim \mathcal{N}(\mu(a_i, k), \sigma(a_i, k))$ where

$$
\begin{aligned}
\mu(a) &= -1.2 + 0.3\log(a) \\
\sigma(a) &= 2.8 - 0.8\log(a) \\
\mu(a, k) &= \begin{cases}
\frac{1}{3}\mu(a) & \text{if } k = 0 \\
\mu(a) & \text{if } k = 1 \\
0.05 + 1.5\mu(a) - 0.01k & \text{if } k \ge 2 \\
\end{cases} \\
\sigma(a, k) &= \begin{cases}
\frac{1}{3}\sigma(a) & \text{if } k = 0 \\
\sigma(a) & \text{if } k = 1 \\
0.01 + \sigma(a) + 0.01k & \text{if } k \ge 2 \\
\end{cases}
\end{aligned}
$$

### 4. Generate Outcome $y_{it}$

$$
y_{it} = \alpha_i + \lambda_t + \tau_{i, k} + \varepsilon_{it}
$$

where $\varepsilon_{it} \sim \mathcal{N}(0, 025)$.

## Simulation

To see its estimation properties,
I simulate the data set 100 times and compare the true values with
the mean, 2.5% quantile, and 97.5% quantile of the estimated values.

```{r simulation}
set.seed(1234)
n_sim <- 100

sim <- function() {
  mdl_base <- gen_data() |>
    indcp(yname = "y",
          iname = "id",
          tname = "year",
          ename = "cyear",
          bname = "byear",
          compute_var = TRUE)

  summary(mdl_base, agg = "event_age")
}

path_sim <- here::here("vignettes/base_indcp/simulated.rds")

if (!file.exists(path_sim)) {
  simulated <- map(1:n_sim, ~sim(), .progress = TRUE) |>
    list_rbind()
  save(simulated, file = path_sim)
} else {
  load(path_sim)
}
```

```{r}
val_true <- tibble(
  event_age = 25:34,
  tau1 = -1.2 + 0.3 * log(event_age),
  sd1 = 3 - 0.75 * log(event_age)) |>
  slice(rep(1:n(), each = 6)) |>
  mutate(rel_time = rep(0:5, 10),
         mean = case_when(
           rel_time == 0 ~ tau1 / 3,
           rel_time == 1 ~ tau1,
           rel_time > 1 ~ 0.05 + 1.5 * tau1 - 0.02 * rel_time),
         sd = case_when(
            rel_time == 0 ~ sd1 / 3,
            rel_time == 1 ~ sd1,
            rel_time > 1 ~ 0.01 + sd1 + 0.02 * rel_time),
         lbl = "Truth") |>
  select(-c(tau1, sd1))

df_plot <- simulated |>
  mutate(sd = sqrt(var)) |>
  summarize(across(c(mean, sd), list("lci" = ~quantile(.x, 0.025),
                                     "rci" = ~quantile(.x, 0.975))),
            across(c(mean, sd), mean),
            .by = c(event_age, rel_time)) |>
  mutate(lbl = "Simulation") |>
  bind_rows(val_true) |>
  mutate(lbl_facet = paste0("Age ", event_age)) |>
  filter(between(event_age, 25, 34))
```

### Mean of $\tau_{ik}$

```{r, fig.align='center', fig.height=4.33, fig.width=7}
df_plot |>
  ggplot(aes(x = rel_time, y = mean, ymin = mean_lci, ymax = mean_rci,
           color = lbl, shape = lbl)) +
  geom_point(position = position_dodge(width = 0.3)) +
  geom_errorbar(width = 0.2, position = position_dodge(width = 0.3)) +
  geom_vline(xintercept = -1, linetype = "dashed") +
  geom_hline(yintercept = 0) +
  scale_color_manual(values = c("#009F8C", "#B75C9D")) +
  facet_wrap(~lbl_facet, ncol = 5) +
  labs(x = "Time to Treatment", y = NULL, color = NULL, shape = NULL) +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        legend.position = "bottom")
```

### Standard Deviation of $\tau_{ik}$

```{r, fig.align='center', fig.height=4.33, fig.width=7}
df_plot |>
  ggplot(aes(x = rel_time, y = sd, ymin = sd_lci, ymax = sd_rci,
           color = lbl, shape = lbl)) +
  geom_point(position = position_dodge(width = 0.3)) +
  geom_errorbar(width = 0.2, position = position_dodge(width = 0.3)) +
  geom_vline(xintercept = -1, linetype = "dashed") +
  geom_hline(yintercept = 0) +
  scale_color_manual(values = c("#009F8C", "#B75C9D")) +
  facet_wrap(~lbl_facet, ncol = 5) +
  labs(x = "Time to Treatment", y = NULL, color = NULL, shape = NULL) +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        legend.position = "bottom")
```









